Solar-Electric and Gas Powered, Long-Endurance UAV Sizing via Geometric Programming

### 2017-02-01 gassolarpaper/gassolarpaper.tex
submission draft

 \usepackage{varioref}%  smart page, figure, table, and equation referencing
 \usepackage{wrapfig}%   wrap figures/tables in text (i.e., Di Vinci style)
 \usepackage{threeparttable}% tables with footnotes
 \usepackage{dcolumn}%   decimal-aligned tabular math columns
  \newcolumntype{d}{D{.}{.}{-1}}
 \usepackage{nomencl}%   nomenclature generation via makeindex
  \makeglossary
 \usepackage{amssymb,amsmath}
 \usepackage{subfigure}% subcaptions for subfigures
 \usepackage{subfigmat}% matrices of similar subfigures, aka small mulitples
 \usepackage{fancyvrb}%  extended verbatim environments
 \fvset{fontsize=\footnotesize,xleftmargin=2em}
 \usepackage{lettrine}%  dropped capital letter at beginning of paragraph
%  \usepackage[dvips]{dropping}% alternative dropped capital package
 \usepackage[colorlinks]{hyperref}%  hyperlinks [must be loaded after dropping]
 \usepackage{float}
 \usepackage{longtable,booktabs,tabularx}
 \restylefloat{table}
 \usepackage{graphicx}
 \usepackage{caption}
 \usepackage{siunitx}
 \usepackage{multicol}
 \usepackage{indentfirst}
 \usepackage[labelfont=bf]{caption}
 \usepackage{multirow}
 \usepackage{setspace}
 \doublespacing

@@

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth]{windvsmonth.pdf}
    \caption{\textbf{Variation of wind speed by month at 45 degrees latitude at altitudes of 15,000 ft, 50,000 ft and 60,000 ft.  Lower, middle and upper lines represent 85th, 90th and 95th percentile winds respectively.}}
	\label{f:windvsmonth}
	\end{center}
\end{figure}

@@

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth]{altvswind.pdf}
    \caption{\textbf{Variation of wind speed with altitude at latitudes of 30, 35, and 45 degrees.  Left, middle and right lines represent 85th, 90th and 95th percentile winds respectively. Wind data includes winds during December for the years 2005-2015.}}
	\label{f:altvswind}
	\end{center}
\end{figure}

@@

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth]{latvswind.pdf}
    \caption{\textbf{Variation of wind speed with latitude at altitudes of 15,000 ft, 50,000 ft and 60,000 ft.  Left, middle and right lines represent 85th, 90th and 95th percentile winds respectively. Wind data includes winds during December for the years 2005-2015.}}
	\label{f:latvswind}
	\end{center}
\end{figure}

@@

\begin{figure}[H]
 \begin{subfigmatrix}{2}% number of columns
     \subfigure[Autocorrelation of Boston wind speeds\label{f:bostonwindatuo}]{\includegraphics{Boston_dec_2015auto.pdf}}
     \subfigure[Boston wind speed data\label{f:bostonwind}]{\includegraphics{Boston_dec_2015.pdf}}
 \end{subfigmatrix}
 \caption{\textbf{Wind speed analysis at 60,000 ft over Boston, Massachusetts during December, 2015.}}
 \label{f:bostonwinds}
\end{figure}

\begin{figure}[H]
 \begin{subfigmatrix}{2}% number of columns
     \subfigure[Autocorrelation of Rome wind speeds\label{f:romewindatuo}]{\includegraphics{Rome_dec_2015auto.pdf}}
     \subfigure[Rome wind speed data\label{f:romewind}]{\includegraphics{Rome_dec_2015.pdf}}
 \end{subfigmatrix}
 \caption{\textbf{Wind speed analysis at 60,000 ft over Rome, Italy during December, 2015.}}
 \label{f:romewinds}
\end{figure}

@@

\begin{align} 
\label{e:gpform}
\text{minimize } f_0(\bold{x}) & \nonumber \\
\text{subject to  } f_i(\bold{x}) &\leq 1, i=1,\ldots,m \\
g_i (\bold{x}) &= 1, i = 1,\ldots,p \nonumber 
\end{align}

@@

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth]{windfitl35.pdf}
    \caption{\textbf{Comparison of a fitted equation of wind speed to air density and percentile wind speed~\eqref{e:windfit35}. (Log space RMS error = 0.043.) Wind speed data includes December wind speeds from 2005-2015 at the 35th degree latitude.}}
	\label{f:windfitl35}
	\end{center}
\end{figure}

@@

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.7\textwidth]{lat30.pdf}
    \caption{\textbf{The available solar power per unit area during the winter solstice in the Northern Hemisphere at $\phi=30^{\circ}$. For a solar-electric powered aircraft to be operational for more than 24 hours, the area under the red curve must be greater than the areas shown in blue, magenta and green.}}
	\label{f:lat30}
	\end{center}
\end{figure}

@@

\begin{align}
    (P/S)_{\text{min}} &= \frac{P_{\text{oper}}}{\eta_{\text{solar}} S_{\text{solar}}} \\
    \eta_{\text{motor}} P_{\text{oper}} &\geq P_{\text{shaft}} + P_{\text{avionics}}. \\
\end{align}

@@

\begin{figure}[H]
 \begin{subfigmatrix}{2}% number of columns
     \subfigure[Monomial Approxmiation for $(E/S)_{\text{day}}$\label{f:Benergy}]{\includegraphics{Benergy.pdf}}
     \subfigure[Monomial Approximation for $(E/S)_{\text{twilight}}$\label{f:Cenergy}]{\includegraphics{Cenergy.pdf}}
 \end{subfigmatrix}
 \caption{\textbf{Monomial approximations for $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$ as functions for $(P/S)_{\text{min}}$ at 30 degrees latitude. (Log-space RMS error = 0.007 and 0.0180)}}
 \label{f:energyapprox}
\end{figure}

@@

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth]{JH01polar.pdf}
    \caption{\textbf{The posynomial fit, Equation~\eqref{e:aerodragprof}, of the \emph{JH01} airfoil drag polar is represented as solid lines at different Reynolds numbers.  The open circles are the calculated drag polars from XFOIL. (Log space RMS error~=~0.00489.)}}
	\label{f:JH01polar}
	\end{center}
\end{figure}

% \subsection{Comparison to Existing Platforms}
% 
% By comparing the results of the optimization model to existing gas-powered, long-endurance platforms confidence is gained in the validity of the models.  
% Known pramaters from the Orion, Predator, Global Hawk, Hermes and Penguin (Table~\ref{t:existingparams}) aircraft are inputed to the optimization model.  
% A comparison of the results from the optimization model and the actual aircraft endurance is summarized in table~\ref{t:existingplatforms} and figure~\ref{f:vehiclecomp}. The model is then solved by maximizing endurance.
% 
% \begin{longtable}{lcccc}
%     \caption{Known Parameters of Exisiting Long Endurance Aircraft}\\
%      \toprule
%      \toprule
%      \label{t:existingparams}
%      Aircraft & MTOW [lbf] & Cruise Speed [m/s] & Altitude [ft] & Payload Weight [lbf]\\
%      \midrule
%      Orion\cite{orion}      & 11,200 & 34  & 15,000 & 1,500\\
%      Predator\cite{predator}   & 2,250  & 70  & 25,000 & 450\\
%      Hermes\cite{hermes}     & 2,600  & 70  & 30,000 & 771\\
%      Global Hawk\cite{globalhawk}& 32,250 & 160 & 50,000 & 3,000\\
%      Penguin\cite{penguin}    & 47    & 22   & 10,000 & 22\\
%      \bottomrule
%  \end{longtable}
% 
%  Additionally, the parameters in table~\ref{t:compparams} are set as constants for each comparison optimization solve. 
% 
% \begin{longtable}{ll}
%     \caption{Fixed Optimization Parameters} \\
%      \toprule
%      \toprule
%      Variable &  Value \\
%      \midrule
%      $\eta_{prop}$        & 0.65   \\
%      $A$                 & 25     \\
%      $f_{\text{structural}}$     & 0.35   \\
%      $\text{BSFC}$ [lb/hp/hr]    & 0.6    \\
%      \bottomrule
%      \label{t:compparams}
%  \end{longtable}
% 
%  \begin{figure}[H]
%  \begin{subfigmatrix}{2}% number of columns
%      \subfigure[MTOW Comparison\label{f:mtowvehiclecomp}]{\includegraphics{mtowvehiclecomp.pdf}}
%      \subfigure[Span Comparison\label{f:bvehiclecomp}]{\includegraphics{bvehiclecomp.pdf}}
%  \end{subfigmatrix}
%  \caption{ \textbf{Comparison of optimization results to existing platforms.  Each aircraft's unique parameters from table~\ref{t:existingparams} are set as fixed constants in the optimization model.} }
%  \label{f:vehiclecomp}
% \end{figure}
%  
% \begin{longtable}{lcccc}
% \caption{Optimization Results Comparison} \\
% \label{t:existingplatforms} \\
%      \toprule
%      \toprule
%      Aircraft & Actual Endurance [days] & Predicted Endurance [days] & Actual Span [ft] & Predicted Span [ft]\\
%      \midrule
%      Orion      & 5 & 4.4   & 132 & 146.8\\
%      Predator   & 1  & 1.7  & 55 & 39.0 \\
%      Hermes     & 1.5 & 1.2 & 50  & 47.27\\
%      Global Hawk& 1.5 & 1.0 & 131 & 100.6\\
%      Penguin    & 1 & 1.6   & 10.8  & 15.14\\
%      \bottomrule
%  \end{longtable}
% 
% 
% Because of limited information about these aircraft the optimization predictions are not exact. However, it does validate the proposition that the geometric programming optimzation model is useful for approximating the size and performance of long-endurance aircraft because the results of the optimization are within the same order of manitude as the existing aircarft.

@@

%      \subfigure[Gas Powered\label{f:mtowvslatgassimple}]{\includegraphics{mtowvslatgassimple.pdf}}
%      \subfigure[Solar-Electric Powered\label{f:mtowvslatsolarsimple}]{\includegraphics{mtowvslatsolarsimple.pdf}}

@@

% 	\includegraphics[width=0.6\textwidth]{mtowvsendsimple.pdf}

@@

% \subsection{Solar-electric Powered Trade Space Exploration}
% 
% For a broader picture of the design space the solar optimization is solved multiple times, producing a set of contour plots for various battery energy densities, structural fraction weights, altitudes, and availabilities.  (Figure~\ref{f:solarcontours}) 
% The matrix contour plot map is a useful reference tool designed to give understanding to key trade offs in the solar-electric powered aircraft sizing problem.
% 
%  \begin{figure}[H]
%  \begin{subfigmatrix}{3}% number of columns
%   \subfigure[40th Latitude, 80\% Wind Speed]{\includegraphics{bcontourl40a80.pdf}}
%   \subfigure[40th Latitude, 85\% Wind Speed]{\includegraphics{bcontourl40a85.pdf}}
%   \subfigure[40th Latitude, 90\% Wind Speed]{\includegraphics{bcontourl40a90.pdf}}
%   \subfigure[35th Latitude, 80\% Wind Speed]{\includegraphics{bcontourl35a80.pdf}}
%   \subfigure[35th Latitude, 85\% Wind Speed]{\includegraphics{bcontourl35a85.pdf}}
%   \subfigure[35th Latitude, 90\% Wind Speed]{\includegraphics{bcontourl35a90.pdf}}
%   \subfigure[30th Latitude, 80\% Wind Speed]{\includegraphics{bcontourl30a80.pdf}}
%   \subfigure[30th Latitude, 85\% Wind Speed]{\includegraphics{bcontourl30a85.pdf}}
%   \subfigure[30th Latitude, 90\% Wind Speed]{\includegraphics{bcontourl30a90.pdf}}
%  \end{subfigmatrix}
%  \caption{ \textbf{Matrix of contour plots of battery specific energy and structural weight fraction vs wing span for solar-electric powered aircraft. Each point on the graph corresponds to a unique design.} }
%  \label{f:solarcontours}
% \end{figure}

@@

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth]{arctanfit.pdf}
    \caption{\textbf{Comparison of monomial approximation to Equation~\ref{e:gustalpha} for $V/V_{\text{gust}} \in [0,0.7]$. (Log space RMS error = 0.0246)}}
	\label{f:arctanfit}
	\end{center}
\end{figure}

@@

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.8\textwidth]{capspar2.jpg}
    \caption{\textbf{Cross sectional view of a cap spar.  The two caps are made of carbon fiber, separated by a foam core. A thin shear web is wrapped around the caps and foam to prevent shearing and buckling.}}
	\label{f:capspar}
	\end{center}
\end{figure}

@@

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth]{taildragpolar.pdf}
    \caption{\textbf{Comparison of XFOIL data for NACA airfoils to GP-compatible generated fit,~\eqref{e:taildrag}. (Log space RMS error = 0.0173)}}
	\label{f:taildragpolar}
	\end{center}
\end{figure}

@@

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth]{powervsweightfit.pdf}
    \caption{\textbf{Comparison of power law fit and engine data from University of North Dakota technical report.\cite{gasengine} (Log space RMS error = 0.34)}}
	\label{f:powervsweightfit}
	\end{center}
\end{figure}

@@

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth]{powertobsfcfit.pdf}
    \caption{\textbf{Comparison of GP-compatible approximation to RCV manufacturing data for the DF70 engine.  Function is representative of engine performance for engines of similar size.  (Log-space RMS error is 0.007) }}
	\label{f:powervsweightfit}
	\end{center}
\end{figure}

@@

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth]{mtowvslat.pdf}
    \caption{\textbf{Trade study comparing the lowest possible max take off weight across a range of possible latitude band requirements. Each point is a unique solution from the optimization. Lower, middle and upper curves were evaluated at the 80th, 90th and 95th percentile winds respectively.  The endurance requirement for the gas powered aircraft in this trade study is 7 days. The next integer latitude value for each of the solar-electric curves is infeasible for the given assumptions.}}
    \label{f:latvsmtowtrade}
	\end{center}
\end{figure}

@@

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth]{mtowvsendurance.pdf}
    \caption{\textbf{Trade study of endurance and size for gas powered architecture., Lower, middle and upper curves were evaluated at the 80th, 90th and 95th percentile winds respectively. Max take off weight was minimized for this analysis.}}
	\label{f:spanvsendurance}
	\end{center}
\end{figure}

@@

\begin{figure}[H]
 \begin{subfigmatrix}{2}% number of columns
     \subfigure[Wingspan\label{f:span}]{\includegraphics{solarobjcomp.pdf}}
     \subfigure[Solar Cell Area\label{f:solarcell}]{\includegraphics{solarobjcomp2.pdf}}
 \end{subfigmatrix}
    \caption{\textbf{Solar aircraft sizing study for two different objective functions: wing span $b$, and solar cell area $S_{\text{solar}}$. Lower, middle and upper curves were evaluated at the 80th, 90th and 95th percentile winds respectively.}}
    \label{f:solarobjcomp}
\end{figure}

@@

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth]{windaltoper.pdf}
 \caption{\textbf{Comparison of optimum operating altitude for the solar-electric powered aircraft with and without a detailed structural model.  With only a structural fraction as the structural model the optimum point is at the minimum wind speed.  With a more detailed structural model the optimum operating altitude is not at the minimum wind speed because of variable interaction between wing weight, air density, and wind speed.}}
 \label{f:altoper}
	\end{center}
\end{figure}

@@

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth]{polarmission.pdf}
    \caption{\textbf{Circles and squares represent the optimized drag and lift coefficient values during the course of mission with and without the wind speed constraint. Solid lines are drag polars at various Reynolds numbers.  The maximum $C_L^{1.5}/C_D$ are marked by a plus sign on the polar curve.}}
 \label{f:polarmission}
	\end{center}
\end{figure}

@@

\begin{figure}[H]
 \begin{subfigmatrix}{2}% number of columns
     \subfigure[$C_L=0.0$\label{f:cpmin}]{\includegraphics{cpmin.pdf}}
     \subfigure[$C_L=1.35$\label{f:cpmax}]{\includegraphics{cpmax.pdf}}
 \end{subfigmatrix}
 \caption{\textbf{Pressure coefficient plots at the minimum and maximum expected $C_L$ at Reynolds number, $Re=3e5$.}}
 \label{f:jhcps}
\end{figure}

>>>>

 \usepackage{varioref}%  smart page, figure, table, and equation referencing
 \usepackage{wrapfig}%   wrap figures/tables in text (i.e., Di Vinci style)
 \usepackage{threeparttable}% tables with footnotes
 \usepackage{dcolumn}%   decimal-aligned tabular math columns
  \newcolumntype{d}{D{.}{.}{-1}}
 \usepackage{nomencl}%   nomenclature generation via makeindex
  \makeglossary
 \usepackage{amssymb,amsmath}
 \usepackage{subfigure}% subcaptions for subfigures
 \usepackage{subfigmat}% matrices of similar subfigures, aka small mulitples
 \usepackage{fancyvrb}%  extended verbatim environments
 \fvset{fontsize=\footnotesize,xleftmargin=2em}
 \usepackage{lettrine}%  dropped capital letter at beginning of paragraph
%  \usepackage[dvips]{dropping}% alternative dropped capital package
 \usepackage[colorlinks]{hyperref}%  hyperlinks [must be loaded after dropping]
 \usepackage{float}
 \usepackage{longtable,booktabs,tabularx}
 \restylefloat{table}
 \usepackage{graphicx}
 \usepackage{caption}
 \usepackage{siunitx}
 \usepackage{multicol}
 \usepackage{indentfirst}
 \usepackage[labelfont=bf]{caption}
 \usepackage{multirow}
 \usepackage{setspace}
 % \usepackage[dvipdfmx]{graphicx} 
 % \usepackage{bmpsize}
 % \doublespacing

@@

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth,natwidth=501,natheight=385]{windvsmonth.pdf}
    \caption{\textbf{Variation of wind speed by month at 45 degrees latitude at altitudes of 15,000 ft, 50,000 ft and 60,000 ft.  Lower, middle and upper lines represent 85th, 90th and 95th percentile winds respectively.}}
	\label{f:windvsmonth}
	\end{center}
\end{figure}

@@

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth,natwidth=538,natheight=405]{altvswind.pdf}
    \caption{\textbf{Variation of wind speed with altitude at latitudes of 30, 35, and 45 degrees.  Left, middle and right lines represent 85th, 90th and 95th percentile winds respectively. Wind data includes winds during December for the years 2005-2015.}}
	\label{f:altvswind}
	\end{center}
\end{figure}

@@

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth,natwidth=576,natheight=432]{latvswind.pdf}
    \caption{\textbf{Variation of wind speed with latitude at altitudes of 15,000 ft, 50,000 ft and 60,000 ft.  Left, middle and right lines represent 85th, 90th and 95th percentile winds respectively. Wind data includes winds during December for the years 2005-2015.}}
	\label{f:latvswind}
	\end{center}
\end{figure}

@@

\begin{figure}[H]
 \begin{subfigmatrix}{2}% number of columns
     \subfigure[Autocorrelation of Boston wind speeds\label{f:bostonwindatuo}]{\includegraphics[natwidth=560,natheight=420]{Boston_dec_2015auto.pdf}}
     \subfigure[Boston wind speed data\label{f:bostonwind}]{\includegraphics[natwidth=560,natheight=420]{Boston_dec_2015.pdf}}
 \end{subfigmatrix}
 \caption{\textbf{Wind speed analysis at 60,000 ft over Boston, Massachusetts during December, 2015.}}
 \label{f:bostonwinds}
\end{figure}

\begin{figure}[H]
 \begin{subfigmatrix}{2}% number of columns
     \subfigure[Autocorrelation of Rome wind speeds\label{f:romewindatuo}]{\includegraphics[natwidth=560,natheight=420]{Rome_dec_2015auto.pdf}}
     \subfigure[Rome wind speed data\label{f:romewind}]{\includegraphics[natwidth=560,natheight=420]{Rome_dec_2015.pdf}}
 \end{subfigmatrix}
 \caption{\textbf{Wind speed analysis at 60,000 ft over Rome, Italy during December, 2015.}}
 \label{f:romewinds}
\end{figure}

@@

\begin{align} 
\label{e:gpform}
\text{minimize } f_0(\bold{x}) & \nonumber \\
\text{subject to  } f_i(\bold{x}) &\leq 1, i=1,\ldots,m \\
g_i (\bold{x}) &= 1, i = 1,\ldots,p 
\end{align}

@@

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth,natwidth=511,natheight=405]{windfitl35.pdf}
    \caption{\textbf{Comparison of a fitted equation of wind speed to air density and percentile wind speed~\eqref{e:windfit35}. (Log space RMS error = 0.043.) Wind speed data includes December wind speeds from 2005-2015 at the 35th degree latitude.}}
	\label{f:windfitl35}
	\end{center}
\end{figure}

@@

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.7\textwidth,natwidth=572,natheight=405]{lat30.pdf}
    \caption{\textbf{The available solar power per unit area during the winter solstice in the Northern Hemisphere at $\phi=30^{\circ}$. For a solar-electric powered aircraft to be operational for more than 24 hours, the area under the red curve must be greater than the areas shown in blue, magenta and green.}}
	\label{f:lat30}
	\end{center}
\end{figure}

@@

\begin{align}
    (P/S)_{\text{min}} &= \frac{P_{\text{oper}}}{\eta_{\text{solar}} S_{\text{solar}}} \\
    \eta_{\text{motor}} P_{\text{oper}} &\geq P_{\text{shaft}} + P_{\text{avionics}}.
\end{align}

@@

\begin{figure}[H]
 \begin{subfigmatrix}{2}% number of columns
     \subfigure[Monomial Approxmiation for $(E/S)_{\text{day}}$\label{f:Benergy}]{\includegraphics[natwidth=576,natheight=432]{Benergy.pdf}}
     \subfigure[Monomial Approximation for $(E/S)_{\text{twilight}}$\label{f:Cenergy}]{\includegraphics[natwidth=575,natheight=432]{Cenergy.pdf}}
 \end{subfigmatrix}
 \caption{\textbf{Monomial approximations for $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$ as functions for $(P/S)_{\text{min}}$ at 30 degrees latitude. (Log-space RMS error = 0.007 and 0.0180)}}
 \label{f:energyapprox}
\end{figure}

@@

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth,natwidth=536,natheight=405]{JH01polar.pdf}
    \caption{\textbf{The posynomial fit, Equation~\eqref{e:aerodragprof}, of the \emph{JH01} airfoil drag polar is represented as solid lines at different Reynolds numbers.  The open circles are the calculated drag polars from XFOIL. (Log space RMS error~=~0.00489.)}}
	\label{f:JH01polar}
	\end{center}
\end{figure}

% \subsection{Comparison to Existing Platforms}
% 
% By comparing the results of the optimization model to existing gas-powered, long-endurance platforms confidence is gained in the validity of the models.  
% Known pramaters from the Orion, Predator, Global Hawk, Hermes and Penguin (Table~\ref{t:existingparams}) aircraft are inputed to the optimization model.  
% A comparison of the results from the optimization model and the actual aircraft endurance is summarized in table~\ref{t:existingplatforms} and figure~\ref{f:vehiclecomp}. The model is then solved by maximizing endurance.
% 
% \begin{longtable}{lcccc}
%     \caption{Known Parameters of Exisiting Long Endurance Aircraft}\\
%      \toprule
%      \toprule
%      \label{t:existingparams}
%      Aircraft & MTOW [lbf] & Cruise Speed [m/s] & Altitude [ft] & Payload Weight [lbf]\\
%      \midrule
%      Orion\cite{orion}      & 11,200 & 34  & 15,000 & 1,500\\
%      Predator\cite{predator}   & 2,250  & 70  & 25,000 & 450\\
%      Hermes\cite{hermes}     & 2,600  & 70  & 30,000 & 771\\
%      Global Hawk\cite{globalhawk}& 32,250 & 160 & 50,000 & 3,000\\
%      Penguin\cite{penguin}    & 47    & 22   & 10,000 & 22\\
%      \bottomrule
%  \end{longtable}
% 
%  Additionally, the parameters in table~\ref{t:compparams} are set as constants for each comparison optimization solve. 
% 
% \begin{longtable}{ll}
%     \caption{Fixed Optimization Parameters} \\
%      \toprule
%      \toprule
%      Variable &  Value \\
%      \midrule
%      $\eta_{prop}$        & 0.65   \\
%      $A$                 & 25     \\
%      $f_{\text{structural}}$     & 0.35   \\
%      $\text{BSFC}$ [lb/hp/hr]    & 0.6    \\
%      \bottomrule
%      \label{t:compparams}
%  \end{longtable}
% 
%  \begin{figure}[H]
%  \begin{subfigmatrix}{2}% number of columns
%      \subfigure[MTOW Comparison\label{f:mtowvehiclecomp}]{\includegraphics[natwidth=610,natheight=642]{mtowvehiclecomp.pdf}}
%      \subfigure[Span Comparison\label{f:bvehiclecomp}]{\includegraphics[natwidth=610,natheight=642]{bvehiclecomp.pdf}}
%  \end{subfigmatrix}
%  \caption{ \textbf{Comparison of optimization results to existing platforms.  Each aircraft's unique parameters from table~\ref{t:existingparams} are set as fixed constants in the optimization model.} }
%  \label{f:vehiclecomp}
% \end{figure}
%  
% \begin{longtable}{lcccc}
% \caption{Optimization Results Comparison} \\
% \label{t:existingplatforms} \\
%      \toprule
%      \toprule
%      Aircraft & Actual Endurance [days] & Predicted Endurance [days] & Actual Span [ft] & Predicted Span [ft]\\
%      \midrule
%      Orion      & 5 & 4.4   & 132 & 146.8\\
%      Predator   & 1  & 1.7  & 55 & 39.0 \\
%      Hermes     & 1.5 & 1.2 & 50  & 47.27\\
%      Global Hawk& 1.5 & 1.0 & 131 & 100.6\\
%      Penguin    & 1 & 1.6   & 10.8  & 15.14\\
%      \bottomrule
%  \end{longtable}
% 
% 
% Because of limited information about these aircraft the optimization predictions are not exact. However, it does validate the proposition that the geometric programming optimzation model is useful for approximating the size and performance of long-endurance aircraft because the results of the optimization are within the same order of manitude as the existing aircarft.

@@

%      \subfigure[Gas Powered\label{f:mtowvslatgassimple}]{\includegraphics[natwidth=610,natheight=642]{mtowvslatgassimple.pdf}}
%      \subfigure[Solar-Electric Powered\label{f:mtowvslatsolarsimple}]{\includegraphics[natwidth=610,natheight=642]{mtowvslatsolarsimple.pdf}}

@@

% 	\includegraphics[natwidth=610,natheight=642][width=0.6\textwidth,natwidth=610,natheight=642]{mtowvsendsimple.pdf}

@@

% \subsection{Solar-electric Powered Trade Space Exploration}
% 
% For a broader picture of the design space the solar optimization is solved multiple times, producing a set of contour plots for various battery energy densities, structural fraction weights, altitudes, and availabilities.  (Figure~\ref{f:solarcontours}) 
% The matrix contour plot map is a useful reference tool designed to give understanding to key trade offs in the solar-electric powered aircraft sizing problem.
% 
%  \begin{figure}[H]
%  \begin{subfigmatrix}{3}% number of columns
%   \subfigure[40th Latitude, 80\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl40a80.pdf}}
%   \subfigure[40th Latitude, 85\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl40a85.pdf}}
%   \subfigure[40th Latitude, 90\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl40a90.pdf}}
%   \subfigure[35th Latitude, 80\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl35a80.pdf}}
%   \subfigure[35th Latitude, 85\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl35a85.pdf}}
%   \subfigure[35th Latitude, 90\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl35a90.pdf}}
%   \subfigure[30th Latitude, 80\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl30a80.pdf}}
%   \subfigure[30th Latitude, 85\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl30a85.pdf}}
%   \subfigure[30th Latitude, 90\% Wind Speed]{\includegraphics[natwidth=610,natheight=642]{bcontourl30a90.pdf}}
%  \end{subfigmatrix}
%  \caption{ \textbf{Matrix of contour plots of battery specific energy and structural weight fraction vs wing span for solar-electric powered aircraft. Each point on the graph corresponds to a unique design.} }
%  \label{f:solarcontours}
% \end{figure}

@@

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth,natwidth=517,natheight=408]{arctanfit.pdf}
    \caption{\textbf{Comparison of monomial approximation to Equation~\ref{e:gustalpha} for $V/V_{\text{gust}} \in [0,0.7]$. (Log space RMS error = 0.0246)}}
	\label{f:arctanfit}
	\end{center}
\end{figure}

@@

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.8\textwidth,natwidth=3242,natheight=1195]{capspar2.jpg}
    \caption{\textbf{Cross sectional view of a cap spar.  The two caps are made of carbon fiber, separated by a foam core. A thin shear web is wrapped around the caps and foam to prevent shearing and buckling.}}
	\label{f:capspar}
	\end{center}
\end{figure}

@@

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth,natwidth=538,natheight=405]{taildragpolar.pdf}
    \caption{\textbf{Comparison of XFOIL data for NACA airfoils to GP-compatible generated fit,~\eqref{e:taildrag}. (Log space RMS error = 0.0173)}}
	\label{f:taildragpolar}
	\end{center}
\end{figure}

@@

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth,natwidth=524,natheight=405]{powervsweightfit.pdf}
    \caption{\textbf{Comparison of power law fit and engine data from University of North Dakota technical report.\cite{gasengine} (Log space RMS error = 0.34)}}
	\label{f:powervsweightfit}
	\end{center}
\end{figure}

@@

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth,natwidth=530,natheight=414]{powertobsfcfit.pdf}
    \caption{\textbf{Comparison of GP-compatible approximation to RCV manufacturing data for the DF70 engine.  Function is representative of engine performance for engines of similar size.  (Log-space RMS error is 0.007) }}
	\label{f:powervsweightfit}
	\end{center}
\end{figure}

@@

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth,natwidth=528,natheight=405]{mtowvslat.pdf}
    \caption{\textbf{Trade study comparing the lowest possible max take off weight across a range of possible latitude band requirements. Each point is a unique solution from the optimization. Lower, middle and upper curves were evaluated at the 80th, 90th and 95th percentile winds respectively.  The endurance requirement for the gas powered aircraft in this trade study is 7 days. The next integer latitude value for each of the solar-electric curves is infeasible for the given assumptions.}}
    \label{f:latvsmtowtrade}
	\end{center}
\end{figure}

@@

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth,natwidth=527,natheight=405]{mtowvsendurance.pdf}
    \caption{\textbf{Trade study of endurance and size for gas powered architecture., Lower, middle and upper curves were evaluated at the 80th, 90th and 95th percentile winds respectively. Max take off weight was minimized for this analysis.}}
	\label{f:spanvsendurance}
	\end{center}
\end{figure}

@@

\begin{figure}[H]
 \begin{subfigmatrix}{2}% number of columns
     \subfigure[Wingspan\label{f:span}]{\includegraphics[natwidth=511,natheight=398]{solarobjcomp.pdf}}
     \subfigure[Solar Cell Area\label{f:solarcell}]{\includegraphics[natwidth=514,natheight=405]{solarobjcomp2.pdf}}
 \end{subfigmatrix}
    \caption{\textbf{Solar aircraft sizing study for two different objective functions: wing span $b$, and solar cell area $S_{\text{solar}}$. Lower, middle and upper curves were evaluated at the 80th, 90th and 95th percentile winds respectively.}}
    \label{f:solarobjcomp}
\end{figure}

@@

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth,natwidth=519,natheight=405]{windaltoper.pdf}
 \caption{\textbf{Comparison of optimum operating altitude for the solar-electric powered aircraft with and without a detailed structural model.  With only a structural fraction as the structural model the optimum point is at the minimum wind speed.  With a more detailed structural model the optimum operating altitude is not at the minimum wind speed because of variable interaction between wing weight, air density, and wind speed.}}
 \label{f:altoper}
	\end{center}
\end{figure}

@@

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth,natwidth=555,natheight=415]{polarmission.pdf}
    \caption{\textbf{Circles and squares represent the optimized drag and lift coefficient values during the course of mission with and without the wind speed constraint. Solid lines are drag polars at various Reynolds numbers.  The maximum $C_L^{1.5}/C_D$ are marked by a plus sign on the polar curve.}}
 \label{f:polarmission}
	\end{center}
\end{figure}

@@

\begin{figure}[H]
 \begin{subfigmatrix}{2}% number of columns
     \subfigure[$C_L=0.0$\label{f:cpmin}]{\includegraphics[width=0.35\textwidth,natwidth=612,natheight=792,angle=-90]{cpmin.pdf}}
     \subfigure[$C_L=1.35$\label{f:cpmax}]{\includegraphics[width=0.35\textwidth,natwidth=612,natheight=792,angle=-90]{cpmax.pdf}}
 \end{subfigmatrix}
 \caption{\textbf{Pressure coefficient plots at the minimum and maximum expected $C_L$ at Reynolds number, $Re=3e5$.}}
 \label{f:jhcps}
\end{figure}
